---
title: "Preprocessing Orchestration"
subtitle: "From data download to geocoding"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    theme: flatly
    number_sections: true
vignette: >
  %\VignetteIndexEntry{Preprocessing Orchestration}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = TRUE, warning = TRUE)
```

# Preparation

```{r install_packages, eval=FALSE}
# Install tidyschoolvax package (if not already installed)
# devtools::install_github("ACCIDDA/tidyschoolvax")
```


## Load Packages

```{r load-packages}
# For development (working directly in the repo), use:
#   devtools::load_all()
# For a fully installed package, use:
#   library(tidyschoolvax)
library(tidyschoolvax)
```

## Specify Inputs

Edit the variables in this section to configure the run for a given state and vaccine type.

```{r specify-inputs}
project_root <- normalizePath(here::here(), winslash = "/", mustWork = TRUE)
state <- "md"  # <-- Specify state
state_id <- "md"  # <-- Specify state for geocoding search string.
vaccine_type_to_keep <- "mmr"   # <-- specify what vaccine type you want
current_state <- toupper(state)
state_name <- state.name[match(current_state, state.abb)]  # Full state name (e.g., "North Carolina")
# Read Google API key from the environment (set GOOGLEGEO_API_KEY before knitting,
# e.g. in .Renviron or via Sys.setenv() in an untracked secrets file).
# Never hard-code a key here.

google_api_key <- ""  # <-- Specify Google API key
Sys.setenv(GOOGLEGEO_API_KEY = google_api_key)
```

## New state onboarding (run once, then comment out)

When adding a new state for the first time, uncomment the lines below. They scaffold the state-specific download and cleaning scripts from the
package templates and immediately open them in your editor so you can fill in the CUSTOMIZE sections before running the rest of the pipeline.

`project_root` is passed explicitly so the files are always written relative to the same root used by the rest of this script.  It defaults to getwd()
inside the function if omitted. The namespaced calls below assume the tidyschoolvax package is installed (or loaded via devtools::load_all()).
```{r setup_stateskeletons, eval=FALSE}
download_script <- tidyschoolvax::create_state_download_script(state, project_root = project_root)
file.edit(download_script)

cleaning_script <- tidyschoolvax::create_state_cleaning_script(state, project_root = project_root)
file.edit(cleaning_script)
```

Address source preference controls which data source's address is used as the primary address in the output.

| Value | Description |
|-------|-------------|
| `"greatschools"` | Use GreatSchools address (default) |
| `"doe"` | Use Department of Education address |
| `"other"` | Use additional dataset address (if available for the state) |
| `"kinder"` | Use kindergarten data address where available |

Set to `"kinder"` if the state's kindergarten vaccination data includes reliable school addresses that should take priority over other sources.

```{r addr-source-pref}
addr_source_pref <- "kinder"
```

## Set Up Paths

```{r set-paths}
paths <- tidyschoolvax::setup_paths(project_root = project_root, state = state)
list2env(paths, envir = environment())
```


# Download Vaccination Data and Standardize Column Names

Download kindergarten school-level vaccination data, state school data (DOE,
DOA, etc.), GreatSchools.org data for standardized school names, and VaxView
data.

## Step 1a: State-Agnostic Downloads (GreatSchools.org, CDC VaxView)

```{r step-1a-download-agnostic}
tidyschoolvax::download_agnostic_data(
  state            = state,
  state_name       = state_name,
  greatschools_dir = greatschools_dir,
  vaxview_dir      = vaxview_dir
)
```

## Step 1b: State-Specific Downloads (Kindergarten Vax Data, DOE/DOA School Data)

```{r step-1b-download-state}
cat("Step 1b: Download state-specific data...\n")
source(file.path(state_dir, "01_cleaning/01_download.R"))
```

**Expected outputs:**

- Merged annual kindergarten school-level vax data (`kinder.csv` & `kinder.rds`)
- GreatSchools.org data for state (`greatschools.csv` & `greatschools.rds`)
- State data on all schools – DOE, DOA, etc. (`doe.csv` & `doe.rds`)
- VaxView data

```{r check-outputs-1}
check_expected_files(
  paths = c(
    file.path(kinder_dir, "kinder_dat.csv"),
    file.path(kinder_dir, "kinder_dat.rds"),
    file.path(greatschools_dir, "greatschools_dat.csv"),
    file.path(greatschools_dir, "greatschools_dat.rds"),
    file.path(doe_dir, "doe_dat.csv"),
    file.path(doe_dir, "doe_dat.rds"),
    file.path(vaxview_dir, "ChildVaxView.csv"),
    file.path(vaxview_dir, "SchoolVaxView.csv"),
    file.path(vaxview_dir, "TeenVaxView.csv"),
    file.path(vaxview_dir, "vax_view.parquet")
  ),
  step_name = "01_download"
)
```

# Standardize School Names, Assign Unique IDs, and Geocode

Standardize school names by matching across data sources (DOE, GreatSchools,
kinder), assign unique school and county IDs, and use the Google API to
geocode schools and look up school operational status.

```{r step-2-standardize}
cat("Step 2: Standardize school names, assign unique school and county IDs\n")
school_vax_joined <- tidyschoolvax::standardize_schools(
  state_id         = state_id,
  kinder_dir       = kinder_dir,
  greatschools_dir = greatschools_dir,
  doe_dir          = doe_dir,
  state_school_dir = state_school_dir,
  temp_data_dir    = temp_data_dir,
  state_geo_dir    = state_geo_dir,
  state_dir        = state_dir,
  addr_source_pref = addr_source_pref
)
```

**Expected outputs:**

- Data with addresses, school level, county, standardized names, lat/lon, and
  operational status
- County key
- School key

```{r check-outputs-2}
check_expected_files(
  paths = c(
    file.path(temp_data_dir, "kinder_vaccination_clean_02.csv"),
    file.path(temp_data_dir, "kinder_vaccination_clean_02.rds")
  ),
  step_name = "02_standardize_school"
)
```

# State-Specific Fixes

Apply any state-specific data cleaning rules before final formatting.

```{r step-3-state-cleaning}
cat("Step 3: State-specific data cleaning...\n")
source(file.path(state_dir, "01_cleaning/03_state_cleaning.R"))
```

**Expected outputs:**

- State-specific cleaned data (intermediate file) to be used for Step 4

```{r check-outputs-3}
check_expected_files(
  paths = c(
    file.path(temp_data_dir, "kinder_vaccination_clean_03.rds"),
    file.path(temp_data_dir, "kinder_vaccination_clean_03.csv")
  ),
  step_name = "03_state_cleaning"
)
```

# Final Formatting and DQA Checks

Apply final formatting rules and data quality assurance (DQA) checks to
produce a clean dataset ready for the model.

```{r step-4-final-formatting}
cat("Step 4: Final formatting and DQA checks...\n")
tidyschoolvax::run_final_formatting(
  state                = state,
  vaccine_type_to_keep = vaccine_type_to_keep,
  temp_data_dir        = temp_data_dir,
  clean_data_dir       = clean_data_dir,
  vaxview_dir          = vaxview_dir,
  general_data_dir     = general_data_dir,
  state_dir            = state_dir,
  outputs_data_dir     = outputs_data_dir
)
```

**Expected outputs:**

- Cleaned vaccination data with DQA checks applied, ready for geocoding
  (`cleaned_data.rds`, `cleaned_data.csv`)
- DQA `.csv` files with issue cases in the temp directory

```{r check-outputs-4}
check_expected_files(
  paths = c(
    file.path(clean_data_dir, "cleaned_data.rds"),
    file.path(clean_data_dir, "cleaned_data.csv")
  ),
  step_name = "04_final_formatting"
)
```
